Quantifying maternal reproductive output of chondrichthyan fishes

### 2024-05-31 rmds/figures.Rmd
revisions

Figure 3. Accuracy (per cent absolute error) in parameter estimates of $\hat{P_{Max}}$ for 3PLF methods. Large sample sizes were needed to accurately estimate $\hat{P_{Max}}$ and accuracy decreased as the duration of the reproductive cycle increased. Each point reflects a mean value from 300 simulated data sets. Point size denotes mean number of females in maternal condition at a given sample size. \
\
\newpage

@@

Figure 4. Confidence interval coverage for $\hat{L'_{50}}$ for 3PLF methods (high selectivity scenarios). Figure shows the percentage of simulations (n = 300) where the true parameter value fell within the 50% bootstrap confidence interval. Point size denotes mean number of females in maternal condition at a given sample size. \
\
\newpage

>>>>

Figure 3. Accuracy (per cent absolute error) in parameter estimates of $\hat{P_{Max}}$ for 3PLF methods with high maternal selectivity. Large sample sizes were needed to accurately estimate $\hat{P_{Max}}$ and accuracy decreased as the duration of the reproductive cycle increased. Each point reflects a mean value from 300 simulated data sets. Point size denotes mean number of females in maternal condition at a given sample size. \
\
\newpage

@@

Figure 4. Confidence interval coverage for $\hat{L'_{50}}$ for 3PLF methods with high maternal selectivity. Figure shows the percentage of simulations (n = 300) where the true parameter value fell within the 50% bootstrap confidence interval. Point size denotes mean number of females in maternal condition at a given sample size. \
\
\newpage

### 2024-05-31 rmds/manuscript.Rmd
revisions

For the live-bearing and egg-laying class of chondrichthyan fishes a three parameter logistic 'maternity' function with a variable upper asymptote, $P_{Max}$, can be used to predict the average probability of a female giving birth or laying eggs in a season. Although fundamental to calculating the reproductive capacity of a population, few studies report maternity functions, with maturity functions often used as a proxy. Applying logistic models to simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from maternal data and that accuracy, bias, and confidence interval coverage often improved compared to when a fixed value was used. Maturity parameters could be estimated with greater accuracy; however, substituting them for maternity parameters overestimated lifetime reproductive output. Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models. In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction. This study suggests that for sample sizes \> 100, estimation of $P_{Max}$ is generally preferable to using a fixed value.

@@

Despite an increase in use, only a small fraction of reproductive studies in recent years have reported maternity functions. Classification of maternal condition is more data-intensive than maturity condition, ideally requiring monthly sampling over a year or longer in order to establish the timing and duration of the female ovarian and uterine cycles, particularly if a there is a resting period between pregnancies. Such data can be difficult and costly to collect for sharks and rays, which are often data-poor and sampled opportunistically. The teleost-oriented foundations of fisheries science have also contributed to the general lack of awareness of maternity functions; Walker (2005) is the sole description on this type of analysis for chondrichthyans and there are no primary literature sources that describe specific methods for chondrichthyan fishes.

@@

Generally when using the 3PLF methods model performance declined as the periodicity of the reproductive cycle increased. This is seen in the positive correlation between $\hat{L'_{50}}$ and $\hat{P_{Max}}$, whereby $P_{Max}$ affected the magnitude and direction of bias in these parameters (Figure 2). With lower values of $P_{Max}$ it appeared to be more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote. As true underlying $P_{Max}$ decreased there was also a greater potential for positive bias in $\hat{P_{Max}}$. Additionally, longer reproductive cycles resulted in fewer maternal females, which was associated with larger bias (Figure 2).

@@

Relatively few reproductive biology studies have used maternity functions to model maternal reproductive output in chondrichthyan fishes. Where they have been employed the approach has typically been to use a three parameter logistic function with a fixed, user-defined value for the upper asymptote, $P_{Max}$. Here we show that it is feasible to estimate $P_{Max}$ from maternal data, in turn enabling statistical inferences on reproductive periodicity. Applying 3PLF models with estimated and fixed values of $P_{Max}$ to simulated data showed that accuracy, bias, and confidence interval coverage often improved when $P_{Max}$ was estimated. Using a fixed value for $P_{Max}$ in some cases resulted in lower bias at low sample sizes. Overall, relatively large sample sizes were needed to estimated $P_{Max}$ accurately, particularly in species with longer reproductive cyles. This study also demonstrated that recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if $P_{Max}$ was ignored when estimating maternity parameters. Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology. We conclude by discussing the advantages of adopting this approach to quantifying maternal reproductive output and future directions.

@@

Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$ from data, but also identified situations where it may be preferable to fix $P_{Max}$. Providing clear guidelines on this is difficult as model performance was influenced by life history traits and gear selectivity. In most scenarios, sample sizes of at least 100 were needed for the 3PLF-estimated method to approach or exceed the performance of the 3PLF-fixed method. At sample sizes below this it is therefore advisable to fix $P_{Max}$. Larger sample sizes are necessary before attempting to estimate $P_{Max}$ in species with a triennial or longer reproductive cycle, due to the decline in the performance of the 3PLF-estimated method as $P_{Max}$ decreases.

While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately. Outcomes of the simulation study showed that for non-annual reproductive cycles $\geq100$ and $\geq200$ maternal females were typically needed to estimate $P_{Max}$ with \<10% MARE for school and gummy sharks, respectively. This is a much larger error than is achievable from conventional logistic maturity analysis [@roa_estimation_1999]. In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying $R_0$. Relatively wide confidence intervals were also obtained for $P_{Max}$ in the empirical analysis for sandbar sharks, even with a total sample size of \> 1000 females, of which 206 were in maternal condition. A similar sample size was used by Colonello *et al*. [-@colonello_reproductive_2016] to estimate plausible values of $P_{Max}$ and maternity parameters for south Atlantic spiny dogfish.

These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained [@oddone_size_2010; @tribuzio_life_2012]. For smaller samples sizes, placing a prior or penalty on $P_{Max}$ may also help overcome issues of parameter inaccuracy, enabling existing information about reproductive periodicity to be incorporated from other populations or closely-related species [@10.1371/journal.pone.0246734]. Information on length at maturity, which is likely to be relatively easier to estimate, may also be informative in the development of priors, with $L_{50}$ effectively providing a lower bound for $L'_{50}$.

For particularly data-poor species, or where data on pregnant females are sparse or unattainable, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option. From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging. However, the extent to which maturity parameters can provide a good approximation of maternity parameters may be species-specific, depending on how close $L_{50}$ is to $L'_{50}$. While several studies have shown these parameters to be similar [@walker_spatial_2007; @soto-lopez_reproductive_2018], $L'_{50}$ is more frequently shifted to the right of $L_{50}$ [@braccini_determining_2006; @palacios-hernandez_temporal_2020; @montealegre-quijano_sexual_2014; @colonello_reproductive_2016], and this length difference can equate to one or more years. For example, Harry et al. [-@harry_age_2013] found both spot-tail shark, *C. sorrah*, and Australian blacktip shark, *C. tilstoni*, began reproducing the year after reaching sexual maturity. Similarly, Fujinami et al. [-@fujinami_reproductive_2017] found $A'_{50}$ to be 1.4 years older than $A_{50}$ in blue sharks.

@@

This unexpected result may be related to the gradual attainment of asymptotic maternal status. Maturity in sandbar sharks appears to occur over an extended size range. $L_{5}$ and $L_{95}$ are 146cm and 176cm – a length interval that corresponds to around seven years given the sandbar shark's slow growth rate [@hale_age_2013]. With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently. The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially, although few females appear to attain this size [@hale_age_2013]. Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund [@baremore_reproduction_2012], they also reproduce more frequently. This implies that size selective fishing or the removal of larger and older female sharks could have a disproportionately greater impact on population productivity.

@@

We thank John Carlson and all NMFS staff who were involved in collecting the sandbar shark reproductive data used in this study. We also thank Alex Hesp for valuable comments on an early draft of this paper and generosity in sharing R code.

>>>>

For the live-bearing and egg-laying class of chondrichthyan fishes a three parameter logistic 'maternity' function with a variable upper asymptote, $P_{Max}$, can be used to predict the average probability of a female giving birth or laying eggs in a season. Although fundamental to calculating the reproductive capacity of a population, few studies report maternity functions, with maturity functions often used as a proxy. Applying logistic models to simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from maternal data and that accuracy, bias, and confidence interval coverage often improved compared to when a fixed value was used. However, sample sizes of 100 - 200 maternal females were required for accurate estimation of $P_{Max}$. Maturity parameters could be estimated with greater accuracy; however, substituting them for maternity parameters overestimated lifetime reproductive output. Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models. In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction.

@@

Despite an increase in use, only a small fraction of reproductive studies in recent years have reported maternity functions. Classification of maternal condition is more data-intensive than maturity condition, ideally requiring monthly sampling over a year or longer in order to establish the timing and duration of the female ovarian and uterine cycles, particularly if there is a resting period between pregnancies. Such data can be difficult and costly to collect for sharks and rays, which are often data-poor and sampled opportunistically. The teleost-oriented foundations of fisheries science have also contributed to the general lack of awareness of maternity functions; Walker (2005) is the sole description on this type of analysis for chondrichthyans and there are no primary literature sources that describe specific methods for chondrichthyan fishes.

@@

When using the 3PLF methods model performance typically declined as the periodicity of the reproductive cycle increased. This is seen in the positive correlation between $\hat{L'_{50}}$ and $\hat{P_{Max}}$, whereby true underlying $P_{Max}$ affected the magnitude and direction of bias in these parameters (Figure 2). With lower values of $P_{Max}$ it appeared to be more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote. As true underlying $P_{Max}$ decreased there was also a greater potential for positive bias in $\hat{P_{Max}}$. Additionally, longer reproductive cycles resulted in fewer maternal females, which was associated with larger bias (Figure 2).

@@

Relatively few reproductive biology studies have used maternity functions to model maternal reproductive output in chondrichthyan fishes. Where they have been employed the approach has typically been to use a three parameter logistic function with a fixed, user-defined value for the upper asymptote, $P_{Max}$. Here we show that it is feasible to estimate $P_{Max}$ from maternal data, in turn enabling statistical inferences on reproductive periodicity. Applying 3PLF models with estimated and fixed values of $P_{Max}$ to simulated data showed that accuracy, bias, and confidence interval coverage often improved when $P_{Max}$ was estimated. Using a fixed value for $P_{Max}$ in some cases resulted in lower bias at low sample sizes. Overall, however, relatively large sample sizes were needed to accurately estimated $P_{Max}$, particularly in species with longer reproductive cycles. This study also demonstrated that, in some cases, recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if $P_{Max}$ was ignored when estimating maternity parameters. Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology. We conclude by discussing the advantages of adopting this approach to quantifying maternal reproductive output and future directions.

@@

Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$, but was difficult to do so accurately. For non-annual reproductive cycles, $> 100$ school sharks and $> 200$ gummy sharks in maternal condition were typically needed to estimate $P_{Max}$ with \<10% MARE (Figure 5). This indicates that the both sample sizes and levels of uncertainty are likely to be larger than conventional logistic maturity analysis [@roa_estimation_1999]. In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying $R_0$. Relatively wide confidence intervals were also obtained for $P_{Max}$ in the empirical analysis for sandbar sharks, even with a total sample size of \> 1000 females, of which 206 were in maternal condition. A similar sample size was used by Colonello *et al*. [-@colonello_reproductive_2016] to estimate plausible values of $P_{Max}$ and maternity parameters for south Atlantic spiny dogfish.

The simulation study also also identified some situations where it may be preferable to fix $P_{Max}$, particularly at lower sample sizes ($\le$ 100 individuals total). Providing clear guidelines on this is difficult as model performance was species-specific and influenced by life history and gear selectivity. The success of any analysis using a fixed value also ultimately depends on the analyst choosing a suitable value. The potential to introduce a greater level of bias through the incorrect selection of $P_{Max}$ should also be considered if using this approach [@natanson_presence_2019].

Together, these results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained [@oddone_size_2010; @tribuzio_life_2012]. For smaller samples sizes, placing a prior or penalty on $P_{Max}$ may also help overcome issues of parameter inaccuracy, enabling existing information about reproductive periodicity to be incorporated from other populations or closely-related species [@10.1371/journal.pone.0246734]. Information on length at maturity, which is likely to be relatively easier to estimate, may also be informative in the development of priors, with $L_{50}$ effectively providing a lower bound for $L'_{50}$.

For the numerous data-poor chondrichthyan species, or where data on pregnant females are sparse or unattainable, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option. From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging. However, the extent to which maturity parameters can provide a good approximation of maternity parameters may be species-specific, depending on how close $L_{50}$ is to $L'_{50}$. While several studies have shown these parameters to be similar [@walker_spatial_2007; @soto-lopez_reproductive_2018], $L'_{50}$ is more frequently shifted to the right of $L_{50}$ [@braccini_determining_2006; @palacios-hernandez_temporal_2020; @montealegre-quijano_sexual_2014; @colonello_reproductive_2016], and this length difference can equate to one or more years. For example, Harry et al. [-@harry_age_2013] found both spot-tail shark, *C. sorrah*, and Australian blacktip shark, *C. tilstoni*, began reproducing the year after reaching sexual maturity. Similarly, Fujinami et al. [-@fujinami_reproductive_2017] found $A'_{50}$ to be 1.4 years older than $A_{50}$ in blue sharks.

@@

This unexpected result may be related to the gradual attainment of asymptotic maternal status. Maturity in sandbar sharks appears to occur over an extended size range. $L_{5}$ and $L_{95}$ are 146cm and 176cm – a length interval that corresponds to around seven years given the sandbar shark's slow growth rate [@hale_age_2013]. With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently. The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially, although few females appear to attain this size [@hale_age_2013]. As such, the modal reproductive frequency is still at least triennial. Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund [@baremore_reproduction_2012], they also reproduce more frequently. This implies that size selective fishing or the removal of larger and older female sharks could have a disproportionately greater impact on population productivity.

@@

We thank John Carlson and all NMFS staff who were involved in collecting the sandbar shark reproductive data used in this study. We also thank Alex Hesp for valuable comments on an early draft of this paper and generosity in sharing R code. We are grateful to three anonymous reviewers for their detailed and constructive comments.

### 2024-05-31 rmds/manuscript.tex
revisions

@@

Parameter estimation was generally straightforward; with sample sizes \(> 50\) a convergence success rate of \(\geq 95\%\) was achieved in all scenarios (Figure S1, Table S1). Overall, convergence and parameter boundary failures were most common with the 2PLF-maternity method which essentially involved fitting data to an under-parameterized model where \(P_{Max}\) was always fixed at 1. In addition to convergence failures, 12 simulations also generated insufficient maternal data (zero or one maternal females in the simulated data), and were discarded.

Generally when using the 3PLF methods model performance declined as the periodicity of the reproductive cycle increased. The nature of this problem can be seen in the positive correlation between the \(\hat{L'_{50}}\) and \(\hat{P_{Max}}\) parameters whereby \(P_{Max}\) affects the magnitude and direction of bias in these parameters (Figure 2). With lower values of \(P_{Max}\) it was apparently more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote. As true underlying \(P_{Max}\) decreased there was also a greater potential for positive bias in \(\hat{P_{Max}}\) (Figure 2).

For the 3PLF methods, parameter estimates tended to be biased high at lower sample sizes (Figure S2, Figure S3). In some scenarios, particularly with sample sizes \(\leq\) 250, fixing the asymptote was effective at reducing bias, although for larger sample sizes neither method was clearly preferable (Figure S2, Figure S3). The accuracy of parameter estimates varied considerably across different combinations of variables used in the simulation and among parameters (Figure S4, Figure S5). Again, in some scenarios with smaller sample sizes \(\leq\) 250, fixing the asymptote improved accuracy with the 3PLF methods, although at larger sample sizes better accuracy was achieved by estimating \(P_{Max}\) (Figure 3). Overall, \(P_{Max}\) was comparatively more challenging to estimate accurately than \(L'_{50}\) (Figure S4, Figure S5).

@@

Differences in life history and gear selectivity played an important role in the ability to obtain precise and unbiased maternity parameters. Certain combinations of variables in the simulated data resulted in fewer females in maternal or immature condition making parameter estimation difficult (Figure S8, Figure S9). These effects were not necessarily consistent between species and seemed to reflect differences in the underlying population length structure. For example, in low selectivity scenarios for gummy sharks \textasciitilde60\% of individuals were immature compared to \textasciitilde50\% in the corresponding scenarios for school sharks (Figure S9). As \(P_{Max}\) decreased the proportion of maternal females in the analysis also decreased.

@@

The effect of the different methods in ultimately calculating \(R_0\) varied considerably across the various scenarios. For school sharks, estimating \(P_{Max}\), was the most effective way to minimize bias in most scenarios (Figure 5, Figure S10). Substituting maturity parameters (2PLF-maturity method) led to a median overestimation of \(R_0\) by 3.4\% across all scenarios, while the 2PLF-maternity method led to a median underestimate of 11.1\%. (Table S1, Figure S10). In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of \(R_0\) of just 1\% across all scenarios which outperformed the 3PLF methods in several instances (Figure 5, Figure S10). Using the 2PLF method for gummy sharks led to a median overestimate of \(R_0\) of 22\% (Figure S10). For both species, better accuracy in calculating \(R_0\) was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S11). For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of \(P_{Max}\).

@@

Relatively few reproductive biology studies have used maternity functions to model maternal reproductive output in chondrichthyan fishes. Where they have been employed the approach has typically been to use a three parameter logistic function with a fixed, user-defined value for the upper asymptote, \(P_{Max}\). Here we show that it is feasible to estimate \(P_{Max}\) from maternal data, in turn enabling statistical inferences on reproductive periodicity. Applying 3PLF models with estimated and fixed values of \(P_{Max}\) to simulated data showed that accuracy, bias, and confidence interval coverage often improved when \(P_{Max}\) was estimated. Using a fixed value for \(P_{Max}\) in some cases resulted in lower bias at low sample sizes. This study also demonstrated that recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if \(P_{Max}\) was ignored when estimating maternity parameters. Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology. We conclude by discussing the advantages of adopting this approach to quantifying maternal reproductive output and future directions.

@@

While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately. Outcomes of the simulation study suggested that samples sizes of \(\geq500\) would be needed to estimate \(P_{Max}\) with \textless10\% MARE for a triennially reproducing species. This is a much larger error than is achievable from conventional logistic maturity analysis (Roa et al. 1999). In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying \(R_0\). Relatively wide confidence intervals were also obtained for \(P_{Max}\) in the empirical analysis for sandbar sharks, even with a sample size of \textgreater{} 1000. A similar sample size was used by Colonello \emph{et al}. (2016) to estimate plausible values of \(P_{Max}\) and maternity parameters for south Atlantic spiny dogfish. These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained (Oddone et al. 2010, Tribuzio and Kruse 2012).

In light of the data requirements suggested by this study, placing a penalty function on \(P_{Max}\) or using a the use of a prior distribution

For some species, and data on pregnant females are particularly sparse, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans. From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging. However, the extent to which maturity parameters can provide a good approximation of maternity parameters may be species-specific, depending on how close \(L_{50}\) is to \(L'_{50}\). While several studies have shown these parameters to be similar (Walker 2007, Soto-López et al. 2018), \(L'_{50}\) is more frequently shifted to the right of \(L_{50}\) (Braccini et al. 2006, Montealegre-Quijano et al. 2014, Colonello et al. 2016, Palacios-Hernández et al. 2020), and this length difference can equate to one or more years. For example, Harry et al. (2013) found both spot-tail shark, \emph{C. sorrah}, and Australian blacktip shark, \emph{C. tilstoni}, began reproducing the year after reaching sexual maturity. Similarly, Fujinami et al. (2017) found \(A'_{50}\) to be 1.4 years older than \(A_{50}\) in blue sharks.

@@

Another approach to overcoming the

Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate \(P_{Max}\) from data, but also identified situations where it may be preferable to fix \(P_{Max}\). In most scenarios sample sizes of at least 100 were needed for the 3PLF-estimated method to approach or exceed the performance of the 3PLF-fixed method. At sample sizes below this it is therefore advisable to fix \(P_{Max}\). Larger sample sizes are necessary before attempting to estimate \(P_{Max}\) in species with a triennial or longer reproductive cycle, due to the decline in the performance of the 3PLF-estimated method as \(P_{Max}\) decreases.

The success of any analysis using the 3PLF-fixed or 2PLF-maturity methods, as implemented here, ultimately depends on the analyst choosing a suitable value for \(P_{Max}\). In this study, the value for \(P_{Max}\) used in the fixed analyses was chosen based on the proportion of maternal females observed in the simulated data. The potential to introduce a greater level of bias in the analysis through the incorrect selection of \(P_{Max}\) should also be considered if using this approach.

@@

This unexpected result may be related to the gradual attainment of asymptotic maternal status. Maturity in sandbar sharks appears to occur over an extended size range. \(L_{5}\) and \(L_{95}\) are 146cm and 176cm -- a length interval that corresponds to around seven years given the sandbar shark's slow growth rate (Hale and Baremore 2013). With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently. The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially, although few females appear to attain this size (Hale and Baremore 2013). Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund (Baremore and Hale 2012), they also reproduce more frequently. This implies that size selective fishing or the removal of larger and older female sharks could have a disproportionately greater impact on population productivity.

@@

>>>>

\begin{itemize}
\tightlist
\item
  A sentence with caveat about the need for large sample sizes to produce ``feasible'' results is warranted.
\end{itemize}

@@

Parameter estimation was generally straightforward; with sample sizes \(> 50\) a convergence success rate of \(\geq 95\%\) was achieved in all scenarios (Figure S1, Table S1). Overall, convergence and parameter boundary failures were most common with the 2PLF-maternity method which essentially involved fitting data to an under-parameterized model where \(P_{Max}\) was always fixed at 1. In addition to convergence failures, 12 simulations also generated zero maternal females, and were discarded.

When using the 3PLF methods model performance typically declined as the periodicity of the reproductive cycle increased. This is seen in the positive correlation between \(\hat{L'_{50}}\) and \(\hat{P_{Max}}\), whereby true underlying \(P_{Max}\) affected the magnitude and direction of bias in these parameters (Figure 2). With lower values of \(P_{Max}\) it appeared to be more difficult for the model to separate the point of inflection in the logistic model and the upper asymptote. As true underlying \(P_{Max}\) decreased there was also a greater potential for positive bias in \(\hat{P_{Max}}\). Additionally, longer reproductive cycles resulted in fewer maternal females, which was associated with larger bias (Figure 2).

For the 3PLF methods, parameter estimates tended to be biased high at lower sample sizes (Figure S2, Figure S3). In some scenarios, particularly with sample sizes \(\leq\) 250, fixing the asymptote was effective at reducing bias, although for larger sample sizes neither method was clearly preferable (Figure S2, Figure S3). The accuracy of parameter estimates varied considerably across different combinations of variables used and among parameters (Figure S4, Figure S5). Again, in some scenarios with smaller sample sizes \(\leq\) 250, fixing the asymptote improved accuracy with the 3PLF methods, although at larger sample sizes better accuracy was achieved by estimating \(P_{Max}\) (Figure 3). Overall, \(P_{Max}\) was comparatively more challenging to estimate accurately than \(L'_{50}\) (Figure S4, Figure S5).

@@

Differences in life history and gear selectivity played an important role in the ability to obtain accurate and unbiased maternity parameters. Certain combinations of variables in the simulated data resulted in few females in maternal or immature condition, making parameter estimation difficult (Figure S8, Figure S9). This was particularly the case for \(P_{Max}\) which required a relatively high maternal sample size to enable accurate parameter estimation (Figure 5). These effects were not necessarily consistent between species and seemed to reflect differences in the underlying population length structure. For example, in biennial low selectivity scenarios for gummy sharks \textasciitilde60\% of individuals were immature compared to \textasciitilde50\% in the corresponding scenarios for school sharks (Figure S9).

@@

The effect of the different methods in ultimately calculating \(R_0\) varied considerably across the various scenarios. For school sharks, estimating \(P_{Max}\), was the most effective way to minimize bias in most scenarios (Figure S10). Substituting maturity parameters (2PLF-maturity method) led to a median overestimation of \(R_0\) by 3.4\% across all scenarios, while the 2PLF-maternity method led to a median underestimate of 11.1\%. (Figure S10). In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of \(R_0\) of just 1\% across all scenarios which outperformed the 3PLF methods in several instances (Figure S10). Using the 2PLF method for gummy sharks led to a median overestimate of \(R_0\) of 22\% (Figure S10). For both species, better accuracy in calculating \(R_0\) was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S11). For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of \(P_{Max}\).

@@

Relatively few reproductive biology studies have used maternity functions to model maternal reproductive output in chondrichthyan fishes. Where they have been employed the approach has typically been to use a three parameter logistic function with a fixed, user-defined value for the upper asymptote, \(P_{Max}\). Here we show that it is feasible to estimate \(P_{Max}\) from maternal data, in turn enabling statistical inferences on reproductive periodicity. Applying 3PLF models with estimated and fixed values of \(P_{Max}\) to simulated data showed that accuracy, bias, and confidence interval coverage often improved when \(P_{Max}\) was estimated. Using a fixed value for \(P_{Max}\) in some cases resulted in lower bias at low sample sizes. Overall, however, relatively large sample sizes were needed to accurately estimated \(P_{Max}\), particularly in species with longer reproductive cycles. This study also demonstrated that, in some cases, recruitment can be overestimated when maturity data were used to approximate maternal data and was inaccurate if \(P_{Max}\) was ignored when estimating maternity parameters. Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology. We conclude by discussing the advantages of adopting this approach to quantifying maternal reproductive output and future directions.

@@

Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate \(P_{Max}\), but also identified situations where it may be preferable to fix \(P_{Max}\). While the estimation of maternity parameters in most simulations was often possible, it was difficult to do so accurately. Outcomes of the simulation study showed that, for non-annual reproductive cycles, \(\geq100\) and \(\geq200\) maternal females were typically needed to estimate \(P_{Max}\) with \textless10\% MARE for school and gummy sharks, respectively (Figure 5). This is a much larger error than is achievable from conventional logistic maturity analysis (Roa et al. 1999). In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying \(R_0\). Relatively wide confidence intervals were also obtained for \(P_{Max}\) in the empirical analysis for sandbar sharks, even with a total sample size of \textgreater{} 1000 females, of which 206 were in maternal condition. A similar sample size was used by Colonello \emph{et al}. (2016) to estimate plausible values of \(P_{Max}\) and maternity parameters for south Atlantic spiny dogfish.

Providing clear guidelines on when to fix or estimate \(P_{Max}\) is difficult as model performance was species-specific and influenced by life history traits and gear selectivity. Most instances where the 3PLF-fixed method outperformed the 3PLF-estimated method typically occurred at lower sample sizes. At sample sizes \textless100 it is therefore advisable to fix \(P_{Max}\). It is also important to note that accurate estimation of the \(P_{Max}\) parameter required a sample size of 100 - 200 females in maternal condition in simulated data, and that larger sample sizes are needed with longer reproductive cycles.

These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained (Oddone et al. 2010, Tribuzio and Kruse 2012). For smaller samples sizes, placing a prior or penalty on \(P_{Max}\) may also help overcome issues of parameter inaccuracy, enabling existing information about reproductive periodicity to be incorporated from other populations or closely-related species (Smart and Grammer 2021). Information on length at maturity, which is likely to be relatively easier to estimate, may also be informative in the development of priors, with \(L_{50}\) effectively providing a lower bound for \(L'_{50}\).

For the numerous data-poor chondrichthyan species, or where data on pregnant females are sparse or unattainable, the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option. From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging. However, the extent to which maturity parameters can provide a good approximation of maternity parameters may be species-specific, depending on how close \(L_{50}\) is to \(L'_{50}\). While several studies have shown these parameters to be similar (Walker 2007, Soto-López et al. 2018), \(L'_{50}\) is more frequently shifted to the right of \(L_{50}\) (Braccini et al. 2006, Montealegre-Quijano et al. 2014, Colonello et al. 2016, Palacios-Hernández et al. 2020), and this length difference can equate to one or more years. For example, Harry et al. (2013) found both spot-tail shark, \emph{C. sorrah}, and Australian blacktip shark, \emph{C. tilstoni}, began reproducing the year after reaching sexual maturity. Similarly, Fujinami et al. (2017) found \(A'_{50}\) to be 1.4 years older than \(A_{50}\) in blue sharks.

@@

@@

This unexpected result may be related to the gradual attainment of asymptotic maternal status. Maturity in sandbar sharks appears to occur over an extended size range. \(L_{5}\) and \(L_{95}\) are 146cm and 176cm -- a length interval that corresponds to around seven years given the sandbar shark's slow growth rate (Hale and Baremore 2013). With females maturing over a broad range of sizes and ages, it is possible that smaller females reproduce less frequently. The outcomes of this study suggest that larger sharks may ultimately be capable of reproducing biennially, although few females appear to attain this size (Hale and Baremore 2013). As such, the modal reproductive frequency is still at least triennial. Further work would be required to confirm this hypothesis, however, such an outcome would have important implications for management; not only are larger females more fecund (Baremore and Hale 2012), they also reproduce more frequently. This implies that size selective fishing or the removal of larger and older female sharks could have a disproportionately greater impact on population productivity.

@@

\bibitem[\citeproctext]{ref-10.1371ux2fjournal.pone.0246734}
Smart, J.J., and Grammer, G.L. 2021. Modernising fish and shark growth curves with bayesian length-at-age models. Plos One \textbf{16}(2): 1--21. doi: \href{https://doi.org/10.1371/journal.pone.0246734}{10.1371/journal.pone.0246734}.
